---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activities = read.csv("~/work/R/coursera/repdata-012/RepData_PeerAssessment1/activity.csv")
str(activities)
```

## What is mean total number of steps taken per day?

```{r}
df = activities[!is.na(activities$steps),]
any(is.na(df$steps))
```

1. Total number of steps taken each day

```{r}
library(plyr)
t = ddply(df, "date", summarise, total.steps = sum(steps))
hist(t$total.steps, main = "Total number of steps taken each day", xlab = "No. of steps")
```


2. Mean and median total number of steps taken per day

  - mean

```{r}
mean(t$total.steps)
```

  - median

```{r}
median(t$total.steps)
```


## What is the average daily activity pattern?
1.  A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
av = ddply(df, "interval", summarise, mean.steps = mean(steps))
plot(av$interval, av$mean.steps, type = "l",
     main = "Average number of steps taken, averaged across all days",
     xlab = "5-minute interval", ylab = "Average No. of steps")
```


2.  5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps

```{r}
av[av$mean.steps == max(av$mean.steps),]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
