---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activities = read.csv("~/work/R/coursera/repdata-012/RepData_PeerAssessment1/activity.csv")
str(activities)
```

## What is mean total number of steps taken per day?

```{r}
df = activities[!is.na(activities$steps),]
any(is.na(df$steps))
```

1. Total number of steps taken each day

```{r}
library(plyr)
t = ddply(df, "date", summarise, total.steps = sum(steps))
hist(t$total.steps, xlab = "No. of steps", 
     main = "Total number of steps taken each day")
```


2. Mean and median total number of steps taken per day

    +   mean

```{r}
mean(t$total.steps)
```

    +   median

```{r}
median(t$total.steps)
```


## What is the average daily activity pattern?
1.  A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 
    the average number of steps taken, averaged across all days (y-axis)

```{r}
av = ddply(df, "interval", summarise, mean.steps = mean(steps))
plot(av$interval, av$mean.steps, type = "l",
     main = "Average number of steps taken, averaged across all days",
     xlab = "5-minute interval", ylab = "Average No. of steps")
```


2.  5-minute interval, on average across all the days in the dataset, that 
    contains the maximum number of steps

```{r}
av[av$mean.steps == max(av$mean.steps),]
```


## Imputing missing values
1. The total number of missing values in the dataset (i.e. the total number of 
    rows with NAs)
    
```{r}
sum(is.na(activities$steps))
```

2.  Devise a strategy for filling in all of the missing values in the dataset. 
    For example, using *mean* for that day.
    
```{r}
# Fill in missing 'steps' values (NAs) with the mean for that day.
act_processed <- ddply(activities, "date", mutate, 
                       mean.steps = mean(steps, na.rm = TRUE),
                       new.steps = ifelse(is.na(mean.steps), 0, steps))

# It turn out that in the data set either a day has values for all intervals
# or NAs for all intervals. The command below will show this:
#       ddply(activities, "data", summarise, sum(is.na(steps)))

head(act_processed)
act_processed[1000:1010,]
```

3.  Create a new dataset that is equal to the original dataset but with the 
    missing data filled in.

```{r}
activities_new <- activities
activities_new$steps <- act_processed$new.steps
```

4.  Make a histogram of the total number of steps taken each day and Calculate 
    and report the mean and median total number of steps taken per day.

```{r}
t <- ddply(activities_new, "date", summarise, total.steps = sum(steps))
hist(t$total.steps, xlab = "No. of steps", main = "Total number of steps")
mean(t$total.steps)
median(t$total.steps)
```

-   Do these values differ from the estimates from the first part of the assignment? 
    +   Yes

-   What is the impact of imputing missing data on the estimates of the total 
    daily number of steps?
    +   The values of the mean and median total number of steps taken per day 
        callculated are lower because the missing data were filled with values. 
        Hence, this highlights the bias presence of missing data can introduce
        to a calculation.

## Are there differences in activity patterns between weekdays and weekends?
